CLT Results

### 2022-04-02 analysis/clt_irt_analysis.Rmd
add irt analysis and item selection draft

```{r setup, include=FALSE}
library(tidyverse)
library(ggpubr)
library(tidyboot)
library(brms)
library(readxl)
library(ggthemes)

@@

```{r}
icc3 %>% 
  ggplot(aes(x = theta, y = p,group = item, col = aoa_german_comb)) +
  geom_line() +
  geom_hline(yintercept = 0.25, lty = 3, alpha = .75)+
  #facet_wrap(~group)+
  #guides(col = F)+
  scale_color_viridis_c(name = "AoA") +
  labs(title = "ICCs for the 3PL",
       x = expression(theta~('ability on the logit scale')),
       y = expression(italic(p)(y==1))) +
  ylim(0,1)+
  theme_minimal()
```

@@

```{r}
irt3_dif <- brm(
  data = irt_dat,
  family = brmsfamily("bernoulli", "identity"),
  bf(
    correct ~ 0.25 + 0.75 * inv_logit(exp(logalpha) * eta),
    eta ~ 1 + (sex |i| targetWord) + (1 | subjID),
    logalpha ~ 1 + (sex |i| targetWord),
    nl = TRUE
  ),
  prior = prior_va_3pl,
  cores = 6,
  chains = 6,
  iter = 6000
)

@@

@@

```{r}
irt1 <- add_criterion(irt1, "loo", moment_match = TRUE) 
irt2 <- add_criterion(irt2, "loo", moment_match = TRUE) 
irt3 <- add_criterion(irt3, "loo", moment_match = TRUE)

loo_compare(irt1, irt2, irt3)%>%as_tibble(rownames = "model")
```

@@

remove <- c("eye" ,"telephone" ,"strawberry", "plane" ,"snake" ,"fish", "giraffe", "bicycle" ,"belt")

firt_dat <- irt_dat%>%
  filter(!targetWord %in% remove)%>%
  group_by(subjID)%>%
  distinct(targetWord, .keep_all = T)%>%
  pivot_wider(names_from = targetWord, values_from = correct)%>%
  ungroup()%>%
  select(-subjID)

firt_1F <- mirt(firt_dat, 1, itemtype = "2PL",  guess = .25,technical = list(NCYCLES = 1000))
firt_2F <- mirt(firt_dat, 2, itemtype = "2PL",technical = list(NCYCLES = 1000))

@@

firt1 <- mirt(firt_dat, 1, itemtype = "Rasch",  guess = .25)
firt2 <- mirt(firt_dat, 1, itemtype = "2PL",  guess = .25,technical = list(NCYCLES = 1000), SE = T)
firt3 <- mirt(firt_dat, 1, itemtype = "3PL",  guess = .25)

anova(firt1,firt2)

summary(firt_2F)
```

@@

# plot chart
ggplot(format, aes(Theta, P.1, col = aoa_german_comb, group = item)) + 
  geom_line() + 
  #ggtitle('Probability Tracelines') + 
  xlab(expression(theta)) + 
  ylab(expression(P(theta))) + 
  #facet_wrap(~targetWord)+
  ylim(0,1)+
  scale_color_viridis_c(name = "AoA") +
  #geom_hline(aes(yintercept = 0.5)) +
  theme_bw() + 
  theme(
        axis.text.x=element_text(colour="black"),
        axis.text.y=element_text(colour="black"),
        legend.title=element_blank())
```

@@

```
```{r}
params%>%left_join(coefs2PL)%>%
  mutate(slope = ifelse(nlpar == "Easiness", -1,1))%>%
  ggplot(aes(x = Estimate, y = par))+
  geom_abline(aes(slope = slope, intercept = 0), lty = 3, alpha = .5)+
  geom_errorbar(aes(ymin = CI_2.5, ymax = CI_97.5), alpha = .25, col = "dodgerblue")+
  geom_errorbarh(aes(xmin = Q2.5, xmax = Q97.5), alpha = .25, col = "dodgerblue")+
  geom_point(pch = 1, alpha = .75)+
  facet_grid(~nlpar)+
  labs(x = "Bayesian estimate", y = "Frequentist estimate")+
  stat_cor()+
  #ylim(-10,10)+
  #xlim(-10,10)+
  #coord_fixed(ratio = 1)+
  scale_color_viridis_c()+
  theme_bw()
```

>>>>

```{r setup, include=FALSE}
library(tidyverse)
library(ggpubr)
library(tidyboot)
library(brms)
library(readxl)
library(ggthemes)
library(geomtextpath)

@@

```{r}
icc3 %>% 
  ggplot(aes(x = theta, y = p,group = item, col = aoa_german_comb)) +
  geom_line() +
  #geom_textline(aes(label = item)) +
  #geom_texthline(yintercept = 0.25, lty = 3, alpha = .75, label = "guessing",hjust = 0.8)+
  geom_hline(yintercept = 0.25, lty = 3, alpha = .75)+
  #facet_wrap(~group)+
  #guides(col = F)+
  scale_color_viridis_c(name = "AoA") +
  labs(title = "ICCs for the 3PL",
       x = expression(theta~('ability on the logit scale')),
       y = expression(italic(p)(y==1))) +
  ylim(0,1)+
  theme_minimal()
```

```{r}
icc3 %>% 
  ggplot(aes(x = theta, y = p, group = item)) +
  geom_line(data = select(icc3, -item),aes(group = german), alpha = .25, col ="grey") +
  geom_line(col = "firebrick") +
  #geom_textline(aes(label = item)) +
  #geom_texthline(yintercept = 0.25, lty = 3, alpha = .75, label = "guessing",hjust = 0.8)+
  geom_hline(yintercept = 0.25, lty = 3, alpha = .75)+
  facet_wrap(~item)+
  #guides(col = F)+
  scale_color_viridis_d(name = "AoA") +
  labs(title = "ICCs for the 3PL",
       x = expression(theta~('ability on the logit scale')),
       y = expression(italic(p)(y==1))) +
  ylim(0,1)+
  theme_minimal()
```

@@

## Differential Item Functioning

```{r}
irt3_dif <- brm(
  data = irt_dat,
  family = brmsfamily("bernoulli", "identity"),
  bf(
    correct ~ 0.25 + 0.75 * inv_logit(exp(logalpha) * eta),
    eta ~ 1 + (0+ sex |i| targetWord) + (1 | subjID),
    logalpha ~ 1 + (0+ sex |i| targetWord),
    nl = TRUE
  ),
  prior = prior_va_3pl,
  cores = 6,
  chains = 6,
  iter = 6000
)

@@

```{r}
coef_irt3_dif <- coef(irt3_dif)

eta <- bind_rows(
  coef_irt3_dif$targetWord[, , "eta_sexf"] %>%
	as_tibble(rownames = "item")%>%mutate(sex = "female"),
  coef_irt3_dif$targetWord[, , "eta_sexm"] %>%
	as_tibble(rownames = "item")%>%mutate(sex = "male")
  )%>%mutate(nlpar = "Easiness")

alpha <- 
  bind_rows(
    coef_irt3_dif$targetWord[, , "logalpha_sexf"] %>%
	exp() %>%
	as_tibble(rownames = "item")%>%mutate(sex = "female"),
	    coef_irt3_dif$targetWord[, , "logalpha_sexm"] %>%
	exp() %>%
	as_tibble(rownames = "item")%>%mutate(sex = "male")
  )%>%mutate(nlpar = "Discrimination")

params <- bind_rows(eta, alpha) %>%
	select(-Est.Error) %>%
  left_join(aoa%>%rename(item = targetWord))

ggplot(params,aes(reorder(item, -aoa_german_comb), Estimate, ymin = Q2.5, ymax = Q97.5)) +
	facet_wrap("nlpar", scales = "free_x") +
	geom_pointrange(aes(col = sex), position = position_dodge(width = .5)) +
	coord_flip() +
  scale_color_colorblind()+
	labs(x = "Item")+
  theme_minimal()
```

@@

```{r}
irt1 <- add_criterion(irt1, "loo") 
irt2 <- add_criterion(irt2, "loo") 
irt3 <- add_criterion(irt3, "loo")

loo_compare(irt1, irt2, irt3)%>%as_tibble(rownames = "model")

```{r}
irt3_dif <- add_criterion(irt3_dif, "loo")

loo_compare(irt3, irt3_dif)%>%as_tibble(rownames = "model")
```

@@

remove <- c("telephone", "strawberry" ,"plane" ,"snake" ,"fish", "belt")

firt_dat <- irt_dat%>%
  select(-sex, -order)%>%
  filter(!targetWord %in% remove)%>%
  group_by(subjID)%>%
  distinct(targetWord, .keep_all = T)%>%
  pivot_wider(names_from = targetWord, values_from = correct)%>%
  ungroup()%>%
  select(-subjID)

firt_1F <- mirt(firt_dat, 1, itemtype = "2PL",  guess = .25,technical = list(NCYCLES = 5000))
firt_2F <- mirt(firt_dat, 2, itemtype = "2PL",  guess = .25,technical = list(NCYCLES = 5000))

@@

firt1 <- mirt(firt_dat, 1, itemtype = "Rasch",  guess = .25)
firt2 <- mirt(firt_dat, 1, itemtype = "2PL",  guess = .25,technical = list(NCYCLES = 5000), SE = T)
firt3 <- mirt(firt_dat, 1, itemtype = "3PL",  guess = .25)

anova(firt1,firt2)
anova(firt2,firt3)

@@

# plot chart
ggplot(format, aes(Theta, P.1, col = aoa_german_comb, group = item)) + 
  geom_line() + 
  #ggtitle('Probability Tracelines') + 
  xlab(expression(theta)) + 
  ylab(expression(P(theta))) + 
  #facet_wrap(~targetWord)+
  ylim(0,1)+
  scale_color_viridis_c(name = "AoA") +
  geom_hline(yintercept = 0.25, lty = 3, alpha = .75) +
  theme_bw() + 
  theme(
        axis.text.x=element_text(colour="black"),
        axis.text.y=element_text(colour="black"),
        legend.title=element_blank())
```

@@

```
```{r}
params%>%left_join(coefs2PL)%>%
  mutate(slope = ifelse(nlpar == "Easiness", -1,1))%>%
  ggplot(aes(x = Estimate, y = par))+
  geom_abline(aes(slope = slope, intercept = 0), lty = 3, alpha = .5)+
  geom_errorbar(aes(ymin = CI_2.5, ymax = CI_97.5), alpha = .25, col = "dodgerblue")+
  geom_errorbarh(aes(xmin = Q2.5, xmax = Q97.5), alpha = .25, col = "dodgerblue")+
  geom_point(pch = 1, alpha = .75)+
  facet_grid(~nlpar)+
  labs(x = "Bayesian estimate", y = "Frequentist estimate")+
  stat_cor()+
  ylim(-10,10)+
  #xlim(-10,10)+
  #coord_fixed(ratio = 1)+
  scale_color_viridis_c()+
  theme_bw()
```
# Correlation subtest and full test

```{r}
select <-params %>%
  pivot_wider(names_from = nlpar, values_from = Estimate)%>%
  arrange(
    #Easiness, 
    -Discrimination
    )%>%
  slice(1:15)#%>%
  pull("item")

full <- data%>%
  group_by(subjID)%>%
  summarise(mean_full = mean(correct))

sub <- data%>%
  filter(targetWord %in% select)%>%
  group_by(subjID)%>%
  summarise(mean_sub = mean(correct))

comp <- full%>%left_join(sub)

ggplot(comp, aes(x = mean_full, y = mean_sub))+
  geom_abline(intercept = 0, slope = 1, lty = 2, alpha = .75)+
  geom_jitter(pch = 1, alpha = .75, height = 0.01, width = 0.01)+
  stat_cor()+
  ylim(0,1.02)+
  xlim(0,1.02)+
  theme_minimal()

```{r}

dat <- tibble(
  iter = c(),
  items = c(),
  cor = c(),
)

full <- data%>%
  group_by(subjID)%>%
  summarise(mean_full = mean(correct))

for (i in 1:1000) {

  items <-params %>% pull("item")

  sel_items <- sample(items, 15)

  sub <- data%>%
  filter(targetWord %in% sel_items)%>%
  group_by(subjID)%>%
  summarise(mean_sub = mean(correct))

  cor <- cor(full$mean_full,sub$mean_sub)

  df <- tibble(
    iter = i,
    items = sel_items,
    cor = cor
  )

  dat <- bind_rows(dat,df)

}

dat%>%
  group_by(items)%>%
  summarise(mean_cor = mean(cor))%>%
  arrange(-mean_cor)
